---
title: "Curing cancer?"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

At the time I wrote this, you can find the following on the [Wikipedia page for
Burkitt's
Lymphoma](https://en.wikipedia.org/wiki/Burkitt%27s_lymphoma#Prognosis).

> The overall cure rate for Burkitt's lymphoma in developed countries is
> about 90%, but worse in low-income countries. Burkitt's lymphoma is
> uncommon in adults, where it has a worse prognosis (Molyneux et al 2012).
>
> In 2006, treatment with dose-adjusted EPOCH with Rituximab showed
> promising initial results in a small series of patients (n=17), with
> a 100% response rate, and 100% overall survival and progression-free
> survival at 28 months (median follow-up) (Dunleavy et al 2006).

* Molyneux *et al* (2012). Burkitt's Lymphoma.  The Lancet, 379(9822),
  1234-1244.
* Dunleavy *et al* (2006). Novel Treatment of Burkitt Lymphoma with
  Dose-Adjusted EPOCH-Rituximab: Preliminary Results Showing Excellent Outcome.
  Blood, 108(11), 2736–2736.

How likely is it that the Dunleavy 2006 study results, or better, could have
come about by chance?

Let's do a simulation with some random numbers.

Here is one one random number between 1 and 10 (inclusive):

```{r}
patient1 <- sample(1:10, 1)
patient1
```

I can use that to simulate one patient in the trial, with a 90% (default) rate
of cure.  I say that 1 through 9 mean that the patient is cured, and 10 means
that the patient was not cured.

I can ask R to tell me whether this patient is cured (TRUE) or not (FALSE):

```{r}
patient1 < 10
```

Now I ask for 17 patients, the same number as in the trial.

```{r}
patients <- sample(1:10, 17, replace=TRUE)
patients
```

Which were cured?

```{r}
cured <- patients < 10
cured
```

How many were cured? (Count the number of TRUE values)

```{r}
n_cured = sum(cured)
n_cured
```

In this simulation, I find that `r n_cured` patients were cured, in this world,
where each patient has a 90% chance of being cured.

Now we use a small amount of programming to repeat this simulation 10000 times:

```{r}
n_cured_per_trial <- numeric(10000)
for (i in 1:10000) {
  # We repeat everthing inside the curly brackets
  # You'll recognize the logic for a single trial here.
  patients <- sample(1:10, 17, replace=TRUE)
  cured <- patients < 10
  n_cured <- sum(cured)
  # Store the result.
  n_cured_per_trial[i] <- n_cured
}
```

What spread do we get in these simulated numbers of patients?

```{r}
hist(n_cured_per_trial)
```

What's the chance that I'll see all 17 patients being cured, in this simulated
world?

```{r}
# TRUE if number was 17, FALSE otherwise.
all_17 <- n_cured_per_trial == 17
# How many TRUEs?
n_all_17 <- sum(all_17)
# What proportion of TRUEs?
p_all_17 <- n_all_17 / 10000
# Show the result.
p_all_17
```

